---
title: "Colour Howto"
author: "Paul roetman"
date: "11 July 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Judisious use of colour in plots that you are trying to demonstrate can help 
various dimensions "come out" better

* default colours in R are crap.
* recent developments in colours in plots/graphs/etc.
* functions in R and in external packages that are very handy.

Here is an example of a very bad plot.

col = 1 ==> black
col = 2 ==> red
col = 3 ==> green.

```{r}

x <- rnorm(20)
y <- rnorm(20)
df <- data.frame(x=x, y=y)
head(df)

plot(df, col=c(1,2,3), pch=c(19, 20, 21))
```


heat.colors(x) ranges from red (low), through yellow to white (high).
   fire ranges from red through yellow to white.
topo.colors() ranges from blue through green to yellow to white/cyan.
  blue low, green higher, etc.

```{r}
heat.colors(2)
topo.colors(2)
```

## Colour Utilities in R

* *grDevices* package has 2 func
** colorRamp
** colorRampPalette
* both take palettes of colors and help to interpolate between the colors. It kinda blends between the colours.
* *colors()* list names of colours you can use in any plotting function. About 650 of them.

```{r}
head(colors())
tail(colors())
```

### Colour Palette

* *colorRamp* take a palette of colours, and return a function that takes values between 0 and 1, indicating extremes of the colour palette - similar to the *gray* function which does the same for black to white.

return values are a 3 vector of Red Green Blue (RGB)
```{r}
pal <- colorRamp(c("red","blue"))
pal(0)     # red  - one end of the avaialble palette
pal(1)     # blue - other end 
pal(0.5)   # 1/2 red, 1/2 blue -- green!
pal(0.25)  # more red than blue

pal(seq(0,1, len=10))  # fading from red to blue over 10 iterations.
```


* *colorRampPalette* take a palette of colours and return a func that takes an integer arg and returns a vector of colours, interpolating the palette (like heat.colors() and topo.colors())

Still has RGB, but in Hex format.

```{r}
pal2 <- colorRampPalette(c("red", "yellow"))
pal2(2)
print('now select 10, ranging from red to yellow')
pal2(10)
```


## RColorBrewer Package

* one package on CRAN that contains interesting/useful color palettes
* 3 types of palettes
** Sequential
** Diverging
** Qualitative   (not ordered, factors, or categorical)
* Palette info can be used in conjunction with colorRamp() and colorRampPalette()

Colours available:
Top Set: sequential, ordered
middle set: Qualitative: categorical, colours have no relation; but do have a theme.
last set: diverging: 

```{r}
library(RColorBrewer)
cols <- brewer.pal(3, "BuGn")
cols
pal <- colorRampPalette(cols)
image(volcano, col=pal(20))
```

## SmoothScatter

scatter plot of a lot of points (eg 10,000 points).
creates a 2d histogram of points, and plots that.
defaults dark blue to lighter blue.

```{r}
x <- rnorm(10000)
y <- rnorm(10000)
smoothScatter(x,y)
```


## Other Functions

* *rgb* functions produces color via Red Green Blue proportions.
* Colour tranpency added via *alpha* parameter to rgb
* *colorspace* package used for different control over colors.


### alpha example

first plot the info is lost in the centre of the diagram, just a blob.
Second plot info is not lost. can see the more dense regions within the centre.
0.2 is pretty close to zero, so very transperant.

```{r}
x <- rnorm(1000)
y <- rnorm(1000)
par(mfrow=c(1,2))
plot(x,y, col=rgb(0,0,0,1), pch=19)
plot(x,y, col=rgb(0,0,0,0.2), pch=19)
```

## Summary

* careful use of colour in plots/maps/etc make eisier for reader to get what you are trying to say.
* transparency is your friend.

