---
title: "control Howto"
author: "Paul roetman"
date: "04 June 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Control Structures

## IF

if (condition) {
    ## do something
} 
else if (other condition) {
    ## do stuff
} else {
    ## do this if all else fails
}


```{r}
x <- 2
y <- if (x > 3) {
        10
     } else {
        0
     }
y
```


## FOR

```{r}
for (i in 1:10) {
    print(i)
}

x <- 1:5
for (i in seq_along(x)) {
    print(x[i])
}

x <- letters[1:5]
for ( letter in x ) {
    print(letter)
}


```

## WHILE

Randomly pick z until z out of the range of 3 and 10.

```{r}
z <- 5

while ( z >= 3 && z <= 10) {
    print(z)
    count <- rbinom(1,1,0.5)

    if (count == 1) {
        z <- z + 1
    } else {
        z <- z - 1
    }
}

```

## REPEAT

```{r}
x <- 20

repeat {
    print(x)
    if (x > 1) {
        print('going to break as x > 1')
        break
    }
}

```

## NEXT

skip to next iteration.

```{r}
for (i in 1:15) {
   if (i < 10) next
    print(i)
}

```

## FUNCTION

Functions return the value from the LAST command

```{r}
add2 <- function(x,y) {
  x + y
}

```

Return a vector with 10 values:
```{r}
above10 <- function(x) {
	use <- x > 10
	x[use]
}
```
### Function Definitions

Functions are a first class object, treated like any other R object. They can be passed as objects to other functions.

Can have named arguments.
Can have default values for arguments.


Formal Arguments args included in func def.
The formals function returns list of all formal args of a function
function args can be missing or may have default values

args can be matched by position or by name

when evaluating parameters, named args are extracted first, then any other parameters are treated by position. So if a named is first, it is extracted, then the remaining is treated as position 0.

you can (and should) mix positional and named var. eg. args(lm)

args can be partially matched (usually on command line). If there is a partial match that is uniq, it will use it.

default val for a arg is NULL

lazy Eval
args of fuction only evaluated as required.
If second arg is never used, even if not defaulted. does not matter

```{r}
f <- function(a,b) {
    a^2
}
f(2)
```
Note - no errors.

But this will create a warning

f <- function(a,b) {
	print(a)
	print(b)
}
f(45)

#### Special arg "..."
variable number of args, usually passed on to other functions

```{r}
myplot <- function(x,y,type="1", ...) {
    plot(x,y,type=type, ...)
}
```
This will create your own plot arg, with the forced default of type=1. Any other arg is just "passed through"

Also used in OO programming for generic functions. NFI what this means.
eg
    mean
    function(x, .)
    UseMethod("mean")


Can also be used as the FIRST arg; where number of args up front can never be known - eg paste - you can paste together any amount of vars..

```{r}
args(paste) # function(..., sep=" ", collapse = NULL)
```

anything after ... must be named explicitly, and no partial matching.

