---
title: "dates howto.Rmd"
author: "Paul roetman"
date: "13 Jun 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dates

When dealing with dates, need to be in class "Date" or classes "POSIXct","POSIXlt""

## Simple

```{r}
d1 <- date()        # "Tue Jun 13 20:32:24 2017"
d1
class(d1)           # Character
```

## Actual Dates

```{r}
d2 <- Sys.Date()   # "2017-06-13"
d2   
class(d2)          # Date
```

## Format dates

* %d = day as number
* %a = abbr weekday
* %A = full weekday
* %m = month number MM
* %b = abbr month
* %B = full month
* %y = 2 digit year
* %Y = 4 digit year

```{r}
format(d2, '%a %b %d')   # "Tue Jun 13"
```

## Cast string to date with format

```{r}
x <- c("1jan1960", "2jan1972")
z <- as.Date(x, "%d%b%Y")
z
z[1] - z[2]
as.numeric(z[1] - z[2])

```

## Convert to Julian

Julian days is the number of days since the origin.

```{r}
weekdays(d2, abbreviate = FALSE)
weekdays(d2, abbreviate = TRUE)
months(d2,abbreviate = FALSE)
months(d2,abbreviate = TRUE)

julian(d2)
```

## Lubridate

```{r}
library(lubridate)
ymd("20140108")
mdy("08/04/2014")
dmy("01-jan-2014")
ymd_hms("20140108 19:15:01")
ymd_hms("20140108 19:15:01", tz="Pacific/Auckland")

```
Display help for timezone....
```{r}
?Sys.timezone
```

```{r}
x = dmy(c("1jan1960", "2jan1972"))
x
wday(x[1])
wday(x[1], label=TRUE)
```
### Further Reading

http://www.r-statistics.com/2012/03/do-more-with-dates-and-times-in-r-with-lubridate-1-1-0/
lubridate vignette
https://cran.r-project.org/web/packages/lubridate/vignettes/lubridate.html
?POSIXlt


