---
title: "dplry howto"
author: "Paul roetman"
date: "30 May 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# dplyr

## Load

```{r}
library(dplyr)
```

## Version

```{r}
packageVersion("dplyr")
```

## Covert to table format
When printing, output much more informative and compact than standard df.

```{r}
library(reshape2)
head(mtcars)
cran <- tbl_df(mtcars)
```

## SELECT
extract only the cols that we are interested in
Note: cran not modified by this op, a new table is returned with the columns selected

```{r}
select(cran, cyl:hp)   # use range of columns
select(cran, cyl, disp, qsec) # change order of cols and subset.
select(cran, -carb)       # exclude col1
```

## FILTER
extract by row

```{r}
filter(cran, cyl == 6)
filter(cran, cyl == 6, disp==160.0)  # both filters must return true
filter(cran, cyl == 6 | disp==108.0) # OR 
filter(cran, !is.na(cyl))           # remove row where col1 is NA
```

## ARRANGE
rows - order by 

```{r}
c2 <- arrange(cran, cyl)         # order by col1 
c2 <- arrange(cran, desc(cyl))   # order by col1 desc
c2 <- arrange(cran, cyl, disp)   # order by col1, then col2
```

## MUTATE
add columns to table

```{r}
c3 <- mutate(cran, disp, size_mb = disp/2^20, size_g=size_mb/2^10)
```

# SUMMARIZE
collapse dataset to a single row

```{r}
summarize(cran, avg_bytes=mean(disp))
```


## GROUP_BY
summarize data by selection of group

```{r}
c4 <- group_by(cran, disp)    # this adds a "grouping" package to new table
summarize(c4, mean(hp))       # return 1 row per group by col1, with mean(col44)

pack_sum <- summarize(cran,
                      count = n(),
                      unique = n_distinct(cyl),
                      count_of_types_of_cyl = n_distinct(unique),
                      avg_bytes = mean(disp))
```

## CHAINING

 piping commands together!
```{r}

result3 <-
  cran %>%
  group_by(cyl) %>%
  summarize(count = n(),
            unique = n_distinct(cyl),
            count_of_types_of_cyl = n_distinct(unique),
            avg_bytes = mean(disp)
  ) %>%
  filter(unique > 5) %>%
  arrange(desc(cyl), avg_bytes)

# Print result to console
print(result3)
```



  
