---
title: "Graphing - Histogram"
author: "Paul roetman"
date: "02 July 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Histogram

Looks like there are a bunch of graphics parameters that can be added to each graph, see http://stat.ethz.ch/R-manual/R-devel/library/graphics/html/par.html

## Load Data to plot
```{r}

pollution <- read.csv('C:/work/R/datasciencecoursera/Course4_ExploreData/Week1/avgpm25.csv', colClasses = c('numeric','character','factor','numeric','numeric'))
head(pollution)

```

## Histogram Plot

```{r}
hist(pollution$pm25, col = "green")
```

## Histogram Plot with Rug

Can see where the bulk of the data are.

```{r}
hist(pollution$pm25, col = "green")
rug(pollution$pm25)
```

## Change Breaks

```{r}
hist(pollution$pm25, col = "green", breaks = 100)
rug(pollution$pm25)
```

## Vertical lines

lwd = line width

```{r}
hist(pollution$pm25, col = "green", breaks = 100)
abline(v=12, lwd=2)
abline(v=median(pollution$pm25), col= "magenta", lwd = 4)

```


## Multiple Histograms

Note: 
mfrow - A vector of the form c(nr, nc). Subsequent figures will be drawn in an nr-by-nc array on the device by columns (mfcol), or rows (mfrow), respectively.
-- larger numbers reduce the size of the graph (fraction of page - 2 ==> 1/2 page)

mar - A numerical vector of the form c(bottom, left, top, right) which gives the number of lines of margin to be specified on the four sides of the plot. The default is c(5, 4, 4, 2) + 0.1.


```{r}
par(mfrow = c(2,1), mar = c(4,4,2,1))
hist(subset(pollution, region == "east")$pm25, col="green")
hist(subset(pollution, region == "west")$pm25, col="green")
```


## Histogram from base system

```{r}
library(datasets)
hist(airquality$Ozone)
```

Very basic example, with all defaults:
   title
   x and y labels
   

```{r}
x <- rnorm(100)
hist(x)
```


## Histogram from ggplot2

histogram

```{r}
library(ggplot2)
data(mpg)
qplot(hwy, data=mpg, fill=drv)
```
