---
title: "Graphing - all"
author: "Paul roetman"
date: "03 July 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# All graphs

## Print to file

```{r}

pollution <- read.csv('C:/work/R/datasciencecoursera/Course4_ExploreData/Week1/avgpm25.csv', colClasses = c('numeric','character','factor','numeric','numeric'))
head(pollution)

```

## Example points


Display all types of points, runs a graphic program when not run in Rmd files.

```{r}
example(points)
```

## Different file types

pdf: line type graphics, resizes well, portable, not efficient if plot has many objects/points.

svg: XML bases scalable vector graphics - supports animation and interactivity - useful for web based plots

win.metafile: windows meta file

postscript: older format, resizes well, portable, create encapsulated postscript files. Windows does not have postscript viewer (but may be able to print)

## bitmap file types

png: bitmapped format, good for line drawings or images with solid colours. Uses lossless compression

jpg: good for photo and natural scenes, uses lossy compression. Good for plotting many many points, does not resize well.

tiff: creates bitmap image in TIFF format, supports lossless compression (older)

bmp: native windows bitmap format.
