---
title: "Graphing - Scatterplot"
author: "Paul roetman"
date: "02 July 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scatterplot

Looks like there are a bunch of graphics parameters that can be added to each graph, see http://stat.ethz.ch/R-manual/R-devel/library/graphics/html/par.html

## Load Data to plot
```{r}

pollution <- read.csv('C:/users/paul/work/R/data_science_courses/04_ExploratoryAnalysis/exploratoryGraphs/data/avgpm25.csv', colClasses = c('numeric','character','factor','numeric','numeric'))
head(pollution)

```

## Scatterplot

Add a linear model line

lwd: line width
lty: Line Type  (0=blank, 1=solid (default), 2=dashed, 3=dotted, 4=dotdash, 5=longdash, 6=twodash) 

```{r}
with(pollution, plot(latitude, pm25))
abline(h=12, lwd=2, lty=2)
```

## scatterplot with colour

```{r}
with(pollution, plot(latitude, pm25, col = region))
abline(h=12, lwd=2, lty=2)
```

## Mutliple scatterplots


```{r}
par(mfrow=c(1,2), mar=c(5,4,2,1))

with(subset(pollution, region == "west"), plot(latitude, pm25, main = 'West'))
abline(h=12, lwd=2, lty=2)

with(subset(pollution, region == "east"), plot(latitude, pm25, main = 'East'))
abline(h=12, lwd=2, lty=2)
```


## Scatterplot from base system

```{r datasets}
data(cars)
with(cars, plot(speed, dist))

library(datasets)
with(airquality, plot(Wind, Ozone))
title(main = "Ozone and Wind in NY city")   ## add title
```

Can do same thing, but with diff format, and colour some points blue.


```{r}
library(datasets)
with(airquality, plot(Wind, Ozone), main = "Ozone and Wind in NY city")
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue"))

```

Plot everything manually (type = n), and add legend!

```{r}
library(datasets)
with(airquality, plot(Wind, Ozone), main = "Ozone and Wind in NY city", type = "n")
with(subset(airquality, Month == 5), points(Wind, Ozone, col = "blue"))
with(subset(airquality, Month != 5), points(Wind, Ozone, col = "red"))
legend("topright", pch=1, col=c("blue", "red"), legend=c("May", "Not May"))

```

Add a regression line, change plot char

```{r}
library(datasets)
with(airquality, plot(Wind, Ozone), main = "Ozone and Wind in NY city",  pch=25)
model <- lm(Ozone ~ Wind, airquality)
abline(model, lwd=2, col = "blue")
```

### Multiple base plots
```{r}
par(mfrow = c(1,2))
with(airquality, {
        plot(Wind, Ozone, main="Ozone and Wind")
        plot(Solar.R, Ozone, main="Ozone and solar radiation")
})
```

Same, but with a title on the outer box.

```{r}
par(mfrow = c(1,3), mar=c(4,4,2,1), oma=c(0,0,2,0))
with(airquality, {
        plot(Wind, Ozone, main="Ozone and Wind")
        plot(Solar.R, Ozone, main="Ozone and solar radiation")
        plot(Temp, Ozone, main="Ozone and Temperature")
        mtext("The overall Title goes here", outer=TRUE)
})
```

another example:

```{r}
x <- rnorm(100)
y <- rnorm(100)
z <- rpois(100,2)
par(mfrow = c(2,1))
plot(x,y, pch=20)
plot(x,z, pch=19)
```

```{r}
x <- rnorm(100)
y <- rnorm(100)
z <- rpois(100,2)
par(mfrow = c(2,1))
par("mar" )
par(mar = c(2,2,1,1))
plot(x,y, pch=20)
plot(x,z, pch=19)
```

### men and women example

```{r}
x <- rnorm(100)
y <- x + rnorm(100)
g <- gl(2, 50, labels = c("Male", "Female"))
plot(x,y)                                      # cannot see diff between m and f
```

plot with diff color
```{r}
plot(x,y, type="n")
points(x[g=="Male"], y[g=="Male"], col = "green")
points(x[g=="Female"], y[g=="Female"], col = "blue", pch = 19)
```
### Very basic example
default point (pch) = open circle

```{r}
x <- rnorm(100)
y <- rnorm(100)
plot(x,y)
par(mar = c(2,2,2,2))
plot(x,y)                 # note lost x and y labels
plot(x,y, pch=2)          # change to triangles
text(-2, -2, "label")

```

# Scatterplot from Lattice system

```{r}
library(lattice)
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4,1))
```

```{r}
library(datasets)
library(lattice)
## convert to factor
airquality <- transform(airquality, Month = factor(Month))
xyplot(Ozone~Wind|Month, data=airquality, layout=c(5,1))
```

## scatter plot with custom panel

This is the default behaviour:
```{r}
set.seed(10)
x <- rnorm(100)
f <- rep(0:1, each = 50)
y <- x + f * x + rnorm(100, sd=0.5)
f <- factor(f, labels = c("Group 1","Group 2"))
xyplot(y~x|f, layout=c(2,1))   ## plot with 2 panels
```

Now with custom panel
```{r}
xyplot(y~x|f, panel=function(x,y,...){
        panel.xyplot(x,y,...)              ## first call to default panel function for xyplot
        panel.abline(h=median(y), lty=2)
})
```
or with regression line
```{r}
xyplot(y~x|f, panel=function(x,y,...){
        panel.xyplot(x,y,...)              ## first call to default panel function for xyplot
        panel.lmline(x,y,col=2)            ## overlay simple linear regression line
})
```
## Scatterplot from ggplot2 plot
drv, f=front wheel drive, r = rear, 4 = 4wd.

```{r}
library(ggplot2)
data(mpg)
str(mpg)
qplot(displ, hwy, data=mpg)
#    qplot(x coord, y coord, dataframe)
```

### Modify aesthetics

Note the auto legend....

```{r}
library(ggplot2)
data(mpg)
qplot(displ, hwy, data=mpg, color=drv)
```

Add a statistic (summary of data)
stat added here is a "smoother" tech name is somethhing like lowet

Here, adding two geom's, first is point (i.e. see the data), the second is the smoothed out line, include 95% (or maybe 5%)
```{r}
library(ggplot2)
data(mpg)
qplot(displ, hwy, data=mpg, geom=c("point", "smooth"))
```

### Facets (or panels)

for facets= a ~ b
a determines columns
b determines rows

we do . ~ drv    so have a rows of graphs

```{r}
library(ggplot2)
data(mpg)
qplot(displ, hwy, data=mpg, facets = . ~ drv)
```


