---
title: "Hierarchical Clustering Howto"
author: "Paul roetman"
date: "10 July 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Hierarchical Clustering

```{r}
set.seed(1234)
par(mar=c(0,0,0,0))
x <- rnorm(12, mean=rep(1:3, each=4), sd=0.2)
y <- rnorm(12, mean=rep(c(1,2,1), each=4), sd=0.2)
plot(x,y, col="blue", pch=19, cex=2)
text(x+0.05, y+0.05, labels=as.character(1:12))

```

Calc distance between points

distance between point 1 and point 2 is 0.34120511,
distance between point 1 and point 7 is 1.49823399

```{r}
dataFrame <- data.frame(x=x, y=y)
dist(dataFrame)
```

So 5 and 6 are closest; so merge, and create a new point between them.
Repeat for 10/11 (next closest).

The dendogram displays the order of merge

5 and 6, then the result of that with 7, then 8.

```{r}
distxy <- dist(dataFrame)
hClustering <- hclust(distxy)
plot(hClustering)
```

Note: no info about cluster count. Need to cut tree (horizontal line) at a level to give you the count of clusters that you are after.

at 1.0, 3 clusters.
at 2.0, 2 clusters.
at 0.5, about 5 clusters.

## Prettier dendigram

download from site
dis0play of hierarchy cluster with coloured leaf labels.

```{r}
myplclust <- function(hclust, lab=hclust$labels, lab.col=rep(1, length(hclust$labels)),
                  hang=0.1,...) {
        y <- rep(hclust$height, 2)
        x <- as.numeric(hclust$merge)
        y <- y[which(x < 0)]
        x <- x[which(x < 0)]
        x <- abs(x)
        y <- y[order(x)]
        x <- x[order(x)]
        plot(hclust, labels=FALSE, hang=hang, ...)
        text(x=x, y=y[hclust$order] - (max(hclust$height)*hang), labels=lab[hclust$order],
             col=lab.col[hclust$order], srt=90, adj=c(1,0.5), xpd=NA, ...)
}
```

## heatmap()

```{r}
dataFrame <- data.frame(x=x, y=y)
set.seed(143)
dataMatrix <- as.matrix(dataFrame)[sample(1:12), ]
heatmap(dataMatrix)
```

good tutorial on heat maps: http://sebastianraschka.com/Articles/heatmaps_in_r.html#clustering

