---
title: "RE Howto (Regular Expression)"
author: "Paul roetman"
date: "13 Jun 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Regular Expressions

* combination of literals and _metacharacters_
* analagy with nateral language - literal text forms words, metachar define grammer.
* re have rich set of metachar.

define literal:

* search for "exact" word

## RE from the ground up.

1. simplest - literals only. Match occurs if literal exists in text being tested.
2. next - End in word, case insensitive: 

## Metacharacters

^          beginning of line
$          end of line
[]         classes of char
               eg [Bb][Uu]  search for any combo of BU.
^[Bb] dd   Line begins with B or b, then space, then dd
[a-z]      match any lower case
[a-zA-Z]   match any case char
[0-9]      any number
[^?.]$     any line that does _NOT_ end in either ? or .
.          match any char
               eg 9.11  search for 9 then anything then 11.
\.         actual period, not wildcard.               
|          OR
               eg flood|fire    return any line with either flood or fire
               
### Combining MetaChar

^[Gg]ood|[Bb]ad   any line begins with good or anywhere in line "bad", case insensitve on first char.
^([Gg]ood|[Bb]ad)  same as above, but BOTH bad and good have to be at begin of line.
?          optional
              eg george (Ww\.)? bush  search for options W. in george bush name.
              
      not sure if these rules are correct, the () will group, not search.             
(.*)       repeat 0 or more times, ie search for brackets with any char in between.
           star is greedy, will match longest possible
^s(.*?)s$  TURN OFF star greedyness.         
(.+)       repeat 1 or more times, search for brackets with at least 1 char.
[0-9]+ (.*)[0-9]+
           search for at least 1 number, space, brackets with any char, then number.
           lecture says brackets not relevent....


```{r}

search_str <- '[0-9]+ (.*)[0-9]+'
#search_str <- '[0-9]*'
str <- '99 abcd33'

grep(search_str, str)

```

###Even more complex

[Bb]ush( +[^ ]+ +){1,5} debate

  starts with bush or Bush
  
  start
        then at least 1 space
        then anything that is not a space [^ ], at least once
        then a space at least once
  end, this repeated between 1 and 5 times
      ie look for betwen 1 and 5 words!
        
  ends with debate
  
{} does counts
m,n  at least m, not more than n
m    exactly m matches
m,   at least m

### Matched text, repeated strings

()   matched by number, reference \1, \2, etc

 +([a-zA-Z]+) +\1 +
 
    at least 1 space 
    then at least 1 char 
    followed by at least 1 space
    followed by same search string again as (), string must be exactly same
    followed by at least 1 space
    
```{r}
search_str <- ' +([a-zA-Z]+) +\1 +'
search_str <- ' +([a-z]+) +\\1 +'


str <- ' aaabbb aaabbb '

grep(search_str, str, value=TRUE)
```
    
    



## Grep

```{r}
if(!file.exists("./tmp")){dir.create("./tmp")}
fileUrl <- "https://data.baltimorecity.gov/Transportation/Baltimore-Fixed-Speed-Cameras/dz54-2aru?accessType=DOWNLOAD"
fileUrl <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD"
download.file(fileUrl, destfile="./tmp/cam.csv", method="curl")
cameraData <- read.csv("./tmp/cam.csv")
names(cameraData)

file.remove("./tmp/cam.csv")
unlink("./tmp")
```



```{r}

names(cameraData) 
grep("Alameda", cameraData$intersection)  # search for str in data, returns ID of found value
grepl("Alameda", cameraData$intersection) # returns True False
print("summarized table results!")
table(grepl("Alameda", cameraData$intersection))  

grep("Alameda", cameraData$intersection, value=TRUE)  # return values, not vector id
grep("zzzzz", cameraData$intersection)    # returns 0 for not found
length(grep("zzzzz", cameraData$intersection))  # len = 0 where nothing found

```

Remove rows that contain a value
```{r}

head(cameraData[!grepl("Alameda", cameraData$intersection),])
```
