---
title: "read.table() howto"
author: "Paul roetman"
date: "30 May 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# read.table() howto

most common used function for reading data. 

args:

* file: name of file or connection
* heading: T/F header line
* sep: string seperator
* colClasses, char vector of class of each col
* nrows, number of rows in dataset (?? is the same as max lines to read???)
* skip, skip first x lines
* stringAsFactors - should char var be coded as factors

## Warnings

* read help
* calc mem
    * rows * cols * size of each col (number=8)
* set comment.char = "" if no comment lines
* colClasses important for large data sets. If only one colClasses, then assumes all columns same.

## Quick'n'Dirty method

```{r}
myfile <- c('dummy.csv')
df <- data.frame(col1=1:6, col2=11:16, col3=sample(letters,6))
write.csv(df, file=myfile, quote=FALSE)

initial <- read.table(myfile, nrows=3, header=TRUE)
classes <- sapply(initial, class)
classes    # reports incorrect, should be NNNNNNCCC
tabAll <- read.table(myfile, colClasses=classes, , header=TRUE)

if (file.exists(myfile)) file.remove(myfile)
```
