---
title: "Strings Howto"
author: "Paul roetman"
date: "13 Jun 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# String Functions

# Get Data set

```{r}
if(!file.exists("./tmp")){dir.create("./tmp")}
fileUrl <- "https://data.baltimorecity.gov/Transportation/Baltimore-Fixed-Speed-Cameras/dz54-2aru?accessType=DOWNLOAD"
fileUrl <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.csv?accessType=DOWNLOAD"
download.file(fileUrl, destfile="./tmp/cam.csv", method="curl")
cameraData <- read.csv("./tmp/cam.csv")
names(cameraData)

file.remove("./tmp/cam.csv")
unlink("./tmp")
```

## Lower Case column names

To remove typo's, name all columns lower case.

```{r}
names <- tolower(names(cameraData))
names(cameraData)
```

Note: toupper also works.

## strsplit

Good for automagically splitting var names
```{r}
splitnames <- strsplit(names(cameraData), "\\.")
splitnames[[5]]
splitnames[[6]]
```

So, to tidy up the names

```{r}
splitnames <- strsplit(names(cameraData), "\\.")
firstElement <- function(x){x[1]}
newnames <- sapply(splitnames, firstElement)
newnames
names(cameraData) <- newnames
```

## Manipulate Variables
Using the Peer Review Data

```{r}
if(!file.exists("./tmp")){dir.create("./tmp")}

fileUrl <- "http://www.plosone.org/article/info:doi/10.1371/journal.pone.0026895"
fileUrl1 <- "https://dl.dropboxusercontent.com/u/7710864/data/reviews-apr29.csv"
fileUrl2 <- "https://dl.dropboxusercontent.com/u/7710864/data/solutions-apr29.csv"
download.file(fileUrl1, destfile="./tmp/reviews.csv", method="curl")
download.file(fileUrl2, destfile="./tmp/solutions.csv", method="curl")
reviews <- read.csv("./tmp/reviews.csv")
solutions <- read.csv("./tmp/solutions.csv")
head(reviews,2)
head(solutions,2)

file.remove("./tmp/reviews.csv")
file.remove("./tmp/solutions.csv")
unlink("./tmp")
```

### Sub - Substitute characters

substitute characters: from, to, source, ignore.case=FALSE, perl=FALSE, fixed=FALSE, useBytes=FALSE)

Note: only will replace the first occurance. To replace multiple, use gsub

example:
```{r}
names(reviews)
sub("_", "", names(reviews),)
```

gsub example:
```{r}
x <- "this_is_a_long_name"
gsub("_", "", x)
```


### GREP

see re_howto.Rmd for more info!

```{r}

names(cameraData) 
grep("Alameda", cameraData$intersection)  # search for str in data, returns ID of found value
grepl("Alameda", cameraData$intersection) # returns True False
print("summarized table results!")
table(grepl("Alameda", cameraData$intersection))  

grep("Alameda", cameraData$intersection, value=TRUE)  # return values, not vector id
grep("zzzzz", cameraData$intersection)    # returns 0 for not found
length(grep("zzzzz", cameraData$intersection))  # len = 0 where nothing found

```

Remove rows that contain a value
```{r}

head(cameraData[!grepl("Alameda", cameraData$intersection),])
```

## stringr
handy string functions
```{r}

library(stringr)
nchar("Jeffery Leek")
substr("Jeffery Leek",6,7)   # first, last (not first, count)
paste('aa','bb')    # default space seperator
paste('a','b',sep='|')
paste0('aa','bb')       # no sep.
str_trim("  aa   ")       #trim leading and trailing strings.
```

## Notes

Variable Names

* lower case
* descriptive (diagnosis, not dx)
* not duplicated
* no underscores or dots or white spaces  (??? why no underscores????)

variables with char values

* usually made into factor variables (depends on application)
* should be descritive (MALE/FEMALE, not 0/1, T/F, M/F, etc)
