---
title: "Subset Howto"
author: "Paul roetman"
date: "27 May 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subsetting different types of data

## Operators

* "["   returns object of same class, can select more than 1 element
* "[["  extract from list or df, extracts single element
* "$"   extract from list or df, similar to "[["

## Subset Vectors


```{r}
# Setup
my_vector <- c('a','b','c')
names(my_vector) <- c('c1','c2','c3')

# Simple
my_vector[1]
my_vector[1:3]
```

### Logical

```{r}
new <- my_vector[my_vector > 'a']
new
```

## Subset Lists

Create a list
```{r}
my_list <- list(foo = 1:4, bar = 0.6)
my_list
```

single bracket returns a list (same class as original)
```{r}
my_list[1]
my_list["bar"]
```

double bracket and $name return a vector
```{r}
my_list[[1]]
my_list$bar
```


 To return the first and third elements as a list:
```{r}
my_list<- list(foo = 1:4, bar = 0.6, baz = "hi")
my_list[c(1,3)]
```

# Use double bracket to reference list element by variable  
```{r}
name <- "foo"
my_list[[name]]
my_list$name

```

