---
title: "Subset Howto"
author: "Paul roetman"
date: "27 May 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subsetting different types of data

## Operators

* "["   returns object of same class, can select more than 1 element
* "[["  extract from list or df, extracts single element
* "$"   extract from list or df, similar to "[["

## Subset Vectors


```{r}
# Setup
my_vector <- c('a','b','c')
names(my_vector) <- c('c1','c2','c3')

# Simple
my_vector[1]
my_vector[1:3]
```

### Logical

```{r}
new <- my_vector[my_vector > 'a']
new
```

## Subset Lists

Create a list
```{r}
my_list <- list(foo = 1:4, bar = 0.6)
my_list
```

single bracket returns a list (same class as original)
```{r}
my_list[1]
my_list["bar"]
```

double bracket and $name return a vector
```{r}
my_list[[1]]
my_list$bar
```


To return the first and third elements as a list:
```{r}
my_list<- list(foo = 1:4, bar = 0.6, baz = "hi")
my_list[c(1,3)]
```

Use double bracket to reference list element by variable  
```{r}
name <- "foo"
my_list[[name]]
my_list$name            # fails

```

## Subset Lists of List

create list of lists
```{r}
x <- list(a=list(10,12,14), b=c(1,2))
x
```

pull out third element of the first list
```{r}
x[[c(1,3)]]
x[[1]][[3]]
```

## Subset Matrix


Create a matrix
```{r}
x <- matrix(1:6, 2,3)
x
```

subset an element, note that the result is a vector, not a matrix
```{r}
x[1,2]
```

extract a row or col (result is a vector)
```{r}
x[1,]
x[,1]
```


extract a matrix, don't drop the MATRIX type, returns a 1,1 matrix
```{r}
x[1,2, drop=FALSE]
```



