---
title: "Tidyr Howto"
author: "Paul roetman"
date: "31 May 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidyr Howto


## GATHER
takes multiple cols, callapses into key-value pairs (use when have columns that are not variables)

```{r} 
library(tidyr)
grade <- c('A','B','C','D','E')
male <- sample(1:10, 5, replace=TRUE)
female <- sample(1:10, 5, replace=TRUE)

students <- data.frame(grade, male, female)
students

gather(students, sex, count, -grade)

```


Each row of the data now represents exactly one observation, characterized by a unique combination of the grade and sex variables. Each of our variables (grade, sex, and count) occupies exactly one column. Thats tidy data!

## SEPARATE

default sep is something that is not a letter or number (like an underscore). Otherwise specify a regex pattern.

```{r}
grade <- c('A','B','C','D','E')
male_1 <- sample(1:10, 5, replace=TRUE)
female_1 <- sample(1:10, 5, replace=TRUE)
male_2 <- sample(1:10, 5, replace=TRUE)
female_2 <- sample(1:10, 5, replace=TRUE)

students2 <- data.frame(grade, male_1, female_1, male_2, female_2)
students2

res <- gather(students2, sex_class, count, -grade)
res

separate(res, sex_class,c("sex", "class"))
```

## CHAIN commands

Use chaining to simplify. This does the same job as above.

```{r}
students2 %>%
  gather(sex_class, count, -grade) %>%
  separate(sex_class,c("sex", "class")) %>%
  print
```


## SPREAD

spread a key/value across multiple columns
```{r}
names <- c('Sally','Sally','Brian','Brian')
test <-  c('Midterm','Final','Midterm','Final')
class <-  c('Class1','Class1','Class1','Class1')
grade <- sample(LETTERS[1:5],  4, replace=TRUE)

students3 <- data.frame(names, test, class, grade)
students3


students3 %>%
  spread(test ,grade ) %>%
  print
```

  
