---
title: "vector_howto.Rmd"
author: "Paul roetman"
date: "27 May 2017"
output: 
  html_document:
    toc: TRUE
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vectors

* Simple vector creation and naming of the values

```{r}
my_range <- 1:3
my_range
names(my_range) <- c('c1','c2','c3')
my_range
```

* Vectors can contain only 1 type
    * char
    * number  (double precision real)
    * integer (specify L)
    * complex
    * logical
```{r}
x <- c(0.5, 0.6)                  # real
x <- c(TRUE, FALSE)               # boolean
x <- c(T,F)                       # short hand for True False
x <- c(1+0i, 2+4i)                # complex 
x <- c(1L, 2L)                    # integer
x <- vector()                     # empty vector
x <- vector("numeric", length=10) # 10 zeros
x
```
Note: list is special vector, can contain different types.

Don't mix types, R will coerce (badly)

## Referencing vectors

```{r}
x <- c('a','b','c','d','e')
x[1]
x[2:4]
x[x>'c']
x_res<- x>'c'
x[x_res]
```

### Remove missing values

```{r} 
x <- c(1,2,NA,4,NA,5)
x
bad <- is.na(x)
x[!bad]
```

### Complete Cases
Campare 2 vectors/lists/? return only cases where both are not NA.

```{r}
x <- c(1,2,NA,4,NA,5)
y <- c('a','b',NA,'d', NA, NA)
good <- complete.cases(x,y)
good
x[good]
y[good]
```

### Vectorised operations

Operations on vectors occur as you would expect!

```{r}
x <- 1:4; y<- 6:9
x + y
y == 8
x * y
```

Other examples

```{r}
z <- c(1,2,3)
z * 2 + 100     # actual calc: z * c(2,2,2) + c(100,100,100)

 c(1, 2, 3, 4) + c(0,10) #  actual calc:  c(1, 2, 3, 4) + c(0,10,0,10)
