---
title: "Air Pollution Case Study"
output:
  html_document: default
  html_notebook: default
---

## Question

What is the question?

Are air pollution levels now lower than in 1999. (now = 2012)

## Data Set
from US EPA - air pollution data, in particular fine particulate matter (dust).
FM 2.5 may affect people lungs (asthma, etc)

I did not download the full set from the site, just did NY state only.

## Load CSV

```{r}
rm(list=ls())
data99 <- read.csv(file='c:/temp/AQ/ad_viz_plotval_data_1999.csv')
data12 <- read.csv(file='c:/temp/AQ/ad_viz_plotval_data_2012.csv')
# str(data99)
# # display a value with NA to check it has loaded correctly.
# data99[444,]
names(data99) <- make.names(names(data99), unique=TRUE)
names(data12) <- make.names(names(data12), unique=TRUE)
head(data99)
```

If column names were not loaded....

```{r}
cnames <- readLines('c:/temp/AQ/ad_viz_plotval_data_1999.csv',1)
cnames <- gsub('"',"", cnames)
cnames <- strsplit(cnames, ",", fixed=TRUE)
# cnames is a list of lists, so only need the first element.
# if we needed to load column headers could do this:
# names(data99) <- cnames[[1]]

```

## Look at PM2.5 data

```{r}
x0_99 <- data99$DAILY_AQI_VALUE
class(x0_99)
str(x0_99)
summary(x0_99)
# If there were missing values, this would give you the percentage
mean(is.na(x0_99))
```

It's interesting that there are no missing values...so the data must have been "cleaned" already.

in the raw data, for all of USA, there are 11% of values missing.

Missing values can play a different role depending on the question. For some questions, it will not matter, for others, it will be critical.

The download I used was for pm2.5 only, so the missing values were probably discarded.

```{r}
x0_12 <- data12$DAILY_AQI_VALUE
class(x0_12)
str(x0_12)
summary(x0_12)
# If there were missing values, this would give you the percentage
mean(is.na(x0_12))
```

Note: there are lots more obs in 2012 than in 1999 - more air poll monitoring stations have  been created.

## Compare

```{r}
summary(x0_99)
summary(x0_12)
```

## first look at data

```{r}
boxplot(x0_99, x0_12)
```

in whole country box plot not usable, so try log

```{r}
boxplot(log10(x0_99), log10(x0_12))
```

In whole country data, there are some negative values - which will be dud readings. In my data, there are a bunch of zero's that do not plot...will look at the zero's instead of the neg (as I have no neg)

## Zero values

Create logical vector of zero values
```{r}
zero_val_99 <- x0_99 == 0
sum(zero_val_99, na.rm=TRUE)
mean(zero_val_99, na.rm=TRUE)
zero_val_12 <- x0_12 == 0
sum(zero_val_12, na.rm=TRUE)
mean(zero_val_12, na.rm=TRUE)

```

So the proportion of zero values is 0.06% and 0.4% - not much.

May need to investigate when the zero values occur (or neg val occur), is it a time of the year
```{r}
dates_99 <- data99$Date
str(dates_99)
dates_99 <- as.Date(as.character(dates_99), "%m/%d/%Y")
str(dates_99)
hist(dates_99, "month")
hist(dates_99[zero_val_99], "month")
```

```{r}
dates_12 <- data12$Date
str(dates_12)
dates_12 <- as.Date(as.character(dates_12), "%m/%d/%Y")
str(dates_12)
hist(dates_12, "month")
hist(dates_12[zero_val_12], "month")
```


# NEW YORK ONLY

In the lecture, they subset the data to look at NY only now.

```{r}
site0_99 <- unique(subset(data99, STATE_CODE==36, c(COUNTY_CODE, AQS_SITE_ID)))

site0_12 <- unique(subset(data12, STATE_CODE==36, c(COUNTY_CODE, AQS_SITE_ID)))

# Now merge the two columns together - county.site_id



site0_99 <- paste(site0_99[,1], site0_99[,2], sep=".")
site0_12 <- paste(site0_12[,1], site0_12[,2], sep=".")


site0_99
site0_12
```

Which are common to both sets

```{r}
both <- intersect(site0_99,site0_12 )
both
```

Now count obs per site.county

```{r}

data99$county.site <- with(data99, paste(COUNTY_CODE,AQS_SITE_ID, sep="."))
data12$county.site <- with(data12, paste(COUNTY_CODE,AQS_SITE_ID, sep="."))

#now count the val

cnt99 <- subset(data99, STATE_CODE==36 & county.site %in% both)
cnt12 <- subset(data12, STATE_CODE==36 & county.site %in% both)
```

```{r}
sapply(split(cnt99,cnt99$county.site), nrow)
sapply(split(cnt12,cnt12$county.site), nrow)
```

pick (for no particular reason) 63.360632008

```{r}
pm99sub <- subset(cnt99, STATE_CODE==36 & county.site == "63.360632008")
pm12sum <- subset(cnt12, STATE_CODE==36 & county.site == "63.360632008")
```

Obs in pm99sub = 92, pm12sub = 416

## Plot Start of results

Plot the data as a function of time.
time on x axis, pm2.5 on y axis

This will visualize if the levels have gone down.

timeseries
```{r}
dates_pm99sub <- pm99sub$Date
dates_pm99sub <- as.Date(as.character(dates_pm99sub), "%m/%d/%Y")
x1_pm99sub <- pm99sub$Daily.Mean.PM2.5.Concentration
plot(dates_pm99sub, x1_pm99sub)
```
```{r}
dates_pm12sub <- pm12sum$Date
dates_pm12sub <- as.Date(as.character(dates_pm12sub), "%m/%d/%Y")
x1_pm12sub <- pm12sum$Daily.Mean.PM2.5.Concentration
plot(dates_pm12sub, x1_pm12sub)
```


Hard to view sperately, so join.

```{r}
par(mfrow = c(1,2), mar = (c(4,4,2,1)))
plot(dates_pm99sub, x1_pm99sub, pch=20)
abline(h=median(x1_pm99sub))
plot(dates_pm12sub, x1_pm12sub, pch=20)
abline(h=median(x1_pm12sub))

```

The abline on the second graph is higher than first, very misleading, as the scale is different. First is above 10, second is below 10.

Todo this, calculate the range of the X data.

```{r}
rng <- range(x1_pm99sub, x1_pm12sub, na.rm=T)
```

```{r}
par(mfrow = c(1,2), mar = (c(4,4,2,1)))
plot(dates_pm99sub, x1_pm99sub, pch=20, ylim=rng)
abline(h=median(x1_pm99sub))
plot(dates_pm12sub, x1_pm12sub, pch=20, ylim=rng)
abline(h=median(x1_pm12sub))
```

## look at just aveage by state now.

Well, I cannot do that, as I only have one states data...will do it by county instead.

Plot the average by county in 99, and again in 12, then connect the dots.

```{r}
mn99 <- with(data99, tapply(Daily.Mean.PM2.5.Concentration, COUNTY_CODE, mean))
summary(mn99)

mn12 <- with(data12, tapply(Daily.Mean.PM2.5.Concentration, COUNTY_CODE, mean))

summary(mn12)
```

Now put that into a dataframe

```{r}
df99 <- data.frame(state=names(mn99), mean=mn99)
df12 <- data.frame(state=names(mn12), mean=mn12)

mrg <- merge(df99, df12, by="state")
head(mrg)
```

now plot it.

```{r}
par(mfrow=c(1,1))
with(mrg, plot(rep(1999, 16 ), mrg[,2], xlim=c(1998, 2013) ))
with(mrg, points(rep(2012, 16 ), mrg[,3]))
segments(rep(1999, 16), mrg[,2], rep(2012, 16), mrg[,3])

```


