---
title: "Exploratory Notebook - Clustering Case Study"
output: html_notebook
---

# Case study of Clustering

Note: When looking at the data, thing to always remember is "what is the question", only explore avenues that head down the path to the answer to "the question". Sense of what you are looking for, and what are the key priorities; what do you want to get out of your dataset. Rules out certain questions, takes you to the next step.

This will generate a rough cut of the "final draft"

This case study is using smart phones to predict human activity.

Data downloaded from : [UCI Website](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones) and raw data from [raw data](http://archive.ics.uci.edu/ml/machine-learning-databases/00240/).

Download locally to c:\temp\UCI; and unzip.

Need to manipulate the data to get to a format that is usable.

```{r}
raw_data <- "C:/Temp/UCI/UCI HAR Dataset/train"
subjects_file <- paste(raw_data, "subject_train.txt", sep="/")
x_file <- paste(raw_data, "X_train.txt", sep="/")
y_file <- paste(raw_data, "y_train.txt", sep="/")

subjects <- read.delim(subjects_file, header=FALSE)
activities <- read.delim(y_file, header=FALSE)
data <- read.fwf(x_file, rep(16, 561))
```

Grab the column headings from the features.txt file and format them.
```{r}
features_file <- "C:/Temp/UCI/UCI HAR Dataset/features.txt"
features <- read.delim(features_file, header=FALSE, sep=" ")
features$V2 <- gsub("[^[:alnum:]]", "", features$V2)
features$col <- paste(sep='_', features$V1, features$V2)
features_headings <- features$col
```
Generate Activity Labels (eg: translate 1 -> walking)
```{r}
activity_labels_file <- "C:/Temp/UCI/UCI HAR Dataset/activity_labels.txt"
activity_labels <- read.fwf(activity_labels_file, c(1,20))
names(activity_labels) <- c('activityid','activityname')
# remove leading and trailing spaces
activity_labels$activityname <- gsub(" ", "",
                                     activity_labels$activityname)
map <- setNames(activity_labels$activityname,
                activity_labels$activityid)
```


Tidy up data a bit

```{r}
names(data) <- features_headings
final_data <- cbind(subjects, data)
final_data$activity <- activities
final_data$activity_full <- apply(final_data$activity, 2,
                                  function(x) map[as.character(x)])

```

Display the first set of columns interested in, and the break down of the activities done.

```{r}
names(final_data[2:13])
table(final_data$activity_full)
```

Plot average acceleration for the first subject. For rest of case study, going to ingore all other subjects and focus on subject 1

```{r}
par(mfrow=c(1,2), mar=c(5,4,1,1))
final_data <- transform(final_data, activity_full=factor(activity_full))
sub1 <- subset(final_data, V1==1)
plot(sub1[,2], col=sub1$activity_full, ylab=names(sub1)[2])
plot(sub1[,3], col=sub1$activity_full, ylab=names(sub1)[3])
legend("bottomright", legend=unique(sub1$activity_full), col=unique(sub1$activity_full), pch=1)
```

## Clustering based on average acceleration

Not a good result, no clean groups of data - colors jumbled together at bottom.

```{r}
myplclust <- function(hclust, lab=hclust$labels, lab.col=rep(1, length(hclust$labels)),
                  hang=0.1,...) {
        y <- rep(hclust$height, 2)
        x <- as.numeric(hclust$merge)
        y <- y[which(x < 0)]
        x <- x[which(x < 0)]
        x <- abs(x)
        y <- y[order(x)]
        x <- x[order(x)]
        plot(hclust, labels=FALSE, hang=hang, ...)
        text(x=x, y=y[hclust$order] - (max(hclust$height)*hang), labels=lab[hclust$order],
             col=lab.col[hclust$order], srt=90, adj=c(1,0.5), xpd=NA, ...)
}
```

```{r}
distanceMatrix <- dist(sub1[2:4])
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col=unclass(sub1$activity_full))
```

As that did not show up any results, try plotting max acceleration.

For things like laying, standing, sitting - nothing, but for the active stuff, there is much more info.


```{r}
par(mfrow=c(1,2))
plot(sub1[,11], pch=19, col=sub1$activity_full, ylab=names(sub1)[11])
plot(sub1[,12], pch=19, col=sub1$activity_full, ylab=names(sub1)[12])
```

so, show dendogram of that info. Can see two very distinct clusters. on right non-moving activities, and on the left is the active stuff.

```{r}
distanceMatrix <- dist(sub1[11:13])
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col=unclass(sub1$activity_full))
```

## SVD

Try a svd, but get rid of the subject and activity columns.

look at first and second sigular vectors, colour code by activity. Can see similar break down - can divide the active, but not the inactive tasks.

```{r}
sub1_stg <- sub1[, -c(1,563,564)]
svd1 <- svd(scale(sub1[, -c(1,563,564)]))
par(mfrow=c(1,2))
plot(svd1$u[,1], col=sub1$activity_full, pch=19)
plot(svd1$u[,2], col=sub1$activity_full, pch=19)

```

```{r}
names(sub1[1:10])
```

## Find maximum contributer

Lets try to figure out which of the contributures is contributing most to the variations.

```{r}
plot(svd1$v[,2], pch=19)

```

```{r}
maxContrib <- which.max(svd1$v[,2])
distanceMatrix <- dist(sub1[, c(10:12, maxContrib)])
hclustering <- hclust(distanceMatrix)
myplclust(hclustering, lab.col = unclass(sub1$activity_full))
```

What is the max Contributer?

```{r}
# need to increment by 1, and I removed the first column in sub1 to svd1.
names(sub1)[maxContrib+1]

```


## Try a Different clustering method

Need to be careful with kmeans, as can get different result based on the start locations.
Can see results of first clustering is jubled. Clust 1 includes both laying and sitting.

If repeat, will get diferent results.


```{r}
kClust <- kmeans(sub1[, -c(1,563,564)], centers = 6)
table(kClust$cluster, sub1$activity_full)
```

Try 100 iterations, look at best results. Still some of the clusters have multiple values.

```{r}
kClust <- kmeans(sub1[, -c(1,563,564)], centers = 6, nstart=100)
table(kClust$cluster, sub1$activity_full)
```

Take these results, see which values drive the cluster location (ie what determines the centre of the cluster)

for us, cluster two is mainly laying.

High values for first 3, rest are very low

```{r}
plot(kClust$center[2, 1:10], pch=19, ylab="Cluster Centre", xlab="")
```

For the cluster on walking, the driving items are completely different, and spread over quite a few....

```{r}
plot(kClust$center[1, 1:10], pch=19, ylab="Cluster Centre", xlab="")

```


So, this is the "fust cut" of the analysis. 

can clearly seperate out the active activities, but not the sedentary activities.

Next round of investigation would be to only look at sedentary, and see what seperates them. By removing all of the active stuff, the scale of difference would be compeltely different.

